Extended Data

### 2021-09-17 doc/manuscript/Dow_manuscript_Nature.Rmd
discussioon
adjust discussion according to issue #57

raise issue #72

There are two main possibilities, which hold contrasting implications for the response of forest C balance to rising spring temperatures.
First, C that is fixed in late summer or fall may be used for woody growth that is not apparent from diameter measurements in the current year: either used for thickening of cell walls, a process that lags behind stem expansion[@cuny_woody_2015], or saved within trees as non-structural carbohydrates and used towards growth the following year[@trumbore_quantifying_2002;@richardson_seasonal_2013].
Tree-ring evidence from SCBI suggests that both processes may influence the climate sensitivity of annual woody growth, but that this dynamic is complex and unlikely to result in substantial additional growth in the following year [@helcoski_growing_2019].
*(we can easily test this: [GitHub issue 57](https://github.com/EcoClimLab/growth_phenology/issues/57))*
Second, any additional C that is not rapidly respired may be allocated, likely in the following growing season, to plant functions other than stem growth, including respiration, reproduction, foliage, fine roots, or root exudates[@trumbore_quantifying_2002]. 
Stem growth is more sensitive to water deficit than photosynthesis is[@muller_water_2011]. This seems to promote the investment of sequestered carbon into root growth in times of drought stress. Zoehner et al 2021[@zohner_how_2021], found that the proportion of root-to-stem growth increased under spring warming, suggesting that increased ecosystem sequestration resulting from spring warming may be promoting the growth of below-ground biomass. However, this study was done under tightly controlled conditions within potted plants and does not represent the environmental conditions of forests like our study does.
Most of this carbon would have a relatively low residence time in the ecosystem[@trumbore_quantifying_2002], although a small fraction would be incorporated into soil organic carbon pools *with residence times rivaling that of wood [REFS]*.
Notably, both of these involve lags of a year or more and would not be captured in interannual variation in NEE, implying that the long-term effect of lengthened growing season on NEE may not be captured in studies, including this one *(This study only considered a time frame of a year or less?)*, that consider time frames of a year or less[@zhou_explaining_2016; @fu_maximum_2019].
Studies examining long-term trends in growing season length and ecosystem C uptake[@keeling_increased_1996;@dragoni_evidence_2011;@keenan_net_2014] -- as opposed to their interannual variation -- within or including the temperate deciduous biome showed increasing trends in both variables, suggesting that the C not allocated to current year growth may have a fairly long residence time within the ecosystem; however, the amount and fate of this carbon remains unknown.

>>>>

There are two main possibilities, which hold contrasting implications for the response of forest C balance to rising spring temperatures.
First, C that is fixed in late summer or fall may be used for woody growth that is not apparent from diameter measurements in the current year: either used for thickening of cell walls, a process that lags behind stem expansion[@cuny_woody_2015], or saved within trees as non-structural carbohydrates and used towards growth the following year[@trumbore_quantifying_2002;@richardson_seasonal_2013].
We cannot rule out the possibility that more carbon-dense cell walls form in years with warmer springs. (**coauthors, do any of you know literature related to this?**)
However, [extension of our tree-ring analysis](https://github.com/EcoClimLab/growth_phenology/issues/57) showed very little correlation between April $T_{max}$ and growth the following year (**#** of 142 RP and **#** of 66 DP species-site combinations), refuting the latter possibility and suggesting that, barring formation of much denser wood in years with warm springs, little of the extra C that may be fixed in these years goes into wood formation.

A second possibility is that any additional C fixed during years with warm springs that is not rapidly respired may be allocated, potentially in the following growing season, to plant functions other than stem growth, including respiration, reproduction, foliage, fine roots, or root exudates[@trumbore_quantifying_2002]. 
Stem growth is more sensitive to water deficit than is photosynthesis[@muller_water_2011]. [This seems to promote the investment of sequestered carbon into root growth in times of drought stress. Zoehner et al 2021[@zohner_how_2021] found that the proportion of root-to-stem growth increased under spring warming, suggesting that increased ecosystem sequestration resulting from spring warming may be promoting the growth of below-ground biomass. ~~However, this study was done under tightly controlled conditions within potted plants and does not represent the environmental conditions of forests.~~](https://github.com/EcoClimLab/growth_phenology/issues/72)
Most of this carbon would have a relatively low residence time in the ecosystem[@trumbore_quantifying_2002], although a small fraction would be incorporated into soil organic carbon pools *with residence times rivaling that of wood [REFS]*.
Notably, both of these involve lags of a year or more and would not be captured in interannual variation in NEE, implying that the long-term effect of lengthened growing season on NEE may not be captured in studies, including this one *(This study only considered a time frame of a year or less?)*, that consider time frames of a year or less[@zhou_explaining_2016; @fu_maximum_2019].
Studies examining long-term trends in growing season length and ecosystem C uptake[@keeling_increased_1996;@dragoni_evidence_2011;@keenan_net_2014] -- as opposed to their interannual variation -- within or including the temperate deciduous biome showed increasing trends in both variables, suggesting that the C not allocated to current year growth may have a fairly long residence time within the ecosystem; however, the amount and fate of this carbon remains unknown.
